---
title: "synth"
output: html_document
date: "2024-04-06"
---

#Display the data:
```{r}
#Load the data
train_data <- read.table("datasets_tp5/synth_train.txt", header = FALSE)
colnames(train_data) <- c("Y", "X1", "X2")

#Display the data
print(train_data)
```

#Dimension of the data:
```{r}
dim(train_data)
```
```{r}
head(train_data)
```
#Graphical representation of the data:
```{r}
plot(train_data[,2:3], col = train_data[,1], pch = 19, xlab = "X1", ylab = "X2")
legend("topright", legend = c("Classe 1", "Classe 2"), col = 1:2, pch = 19)
```
#Prediction of the pointas (0, 0) and (-2, 2):
```{r}
library(class)

X_train <- train_data[,2:3]
Y_train <- train_data[,1]

new_points <- matrix(c(0, 0, -2, 2), ncol = 2, byrow = TRUE)

predictions <- knn(X_train, new_points, Y_train, k = 15)
predictions
```
#Calculating the emprical errors of k = 15, 10, 5, 3, 1:
```{r}
calc_error_rate <- function(predictions, actual) {
  return (sum(predictions != actual) / length(actual))
}

k_values <- c(15, 10, 5, 3, 1)
for (k in k_values) {
  predictions <- knn(X_train, X_train, Y_train, k = k)
  error_rate <- calc_error_rate(predictions, Y_train)
  cat("k =", k, "Empirical error rate:", error_rate, "\n")
}
```
#Decision Frontiers:
```{r}
x1 <- seq(min(X_train[,1]), max(X_train[,1]), length.out = 100)
x2 <- seq(min(X_train[,2]), max(X_train[,2]), length.out = 100)
grid <- expand.grid(x1, x2)

predictions_grid <- knn(X_train, grid, Y_train, k = 15)

plot(grid, col = ifelse(predictions_grid == 1, "red", "blue"), pch = ".", xlab = "X1", ylab = "X2")
points(X_train, col = ifelse(Y_train == 1, "red", "blue"), pch = 19)
legend("topright", legend = c("Class 1", "Class 2"), col = c("red", "blue"), pch = 19)
```
```{r}
test_data <- read.table("datasets_tp5/synth_test.txt", header = FALSE)

X_test <- test_data[,2:3]
Y_test <- test_data[,1]

predictions_test_k15 <- knn(X_train, X_test, Y_train, k = 15)
error_rate_k15 <- calc_error_rate(predictions_test_k15, Y_test)
cat("The empirical error for k = 15:", error_rate_k15, "\n")

predictions_test_k1 <- knn(X_train, X_test, Y_train, k = 1)
error_rate_k1 <- calc_error_rate(predictions_test_k1, Y_test)
cat("The empirical error for k = 1:", error_rate_k1, "\n")

cat("The empirical error for the training set with k = 15:", error_rate, "\n")
```
#Calculating TVN and TVP:
```{r}
TP <- sum(predictions_test_k15 == 1 & Y_test == 1)
FN <- sum(predictions_test_k15 == 2 & Y_test == 1)
TVP <- TP / (TP + FN)

TN <- sum(predictions_test_k15 == 2 & Y_test == 2)
FP <- sum(predictions_test_k15 == 1 & Y_test == 2)
TVN <- TN / (TN + FP)

cat("True positives (TVP):", TVP, "\n")
cat("True négatives (TVN):", TVN, "\n")
```
#prob_knn function:
```{r}
prob_knn <- function(X_train, Y_train, X_test, k) {
  predictions <- knn(X_train, X_test, Y_train, k = k)
  prob_1 <- numeric(length(predictions))
  
  for (i in 1:length(predictions)) {
    prob_1[i] <- sum(Y_train[predictions[i] == Y_train] == 1) / sum(predictions[i] == Y_train)
  }
  
  return(prob_1)
}

#Probability estimations of P(Y = 1|X = x)
test_probs <- prob_knn(train_data[, 1:2], train_data$Y, test_data[, 1:2], k = 15)

#Cost Matrix
cost_matrix <- matrix(c(0, 3, 1, 0), nrow = 2, byrow = TRUE)
rownames(cost_matrix) <- colnames(cost_matrix) <- c(1, 2)

#Bayes Prediction
predicted_classes <- factor(ifelse(test_probs > 0.5, 1, 2), levels = 1:2)

#Misclassification Cost
misclassification_cost <- sum(ifelse(predicted_classes == test_data$Y, 0,
                                     ifelse(predicted_classes == 1, cost_matrix[2, 1],
                                            cost_matrix[1, 2])))
print(paste("Test_probs :", test_probs))
print(paste("Predicted_classes :", predicted_classes))
print(paste("Misclassification_cost :", misclassification_cost))
```
## Methodological Assessment of the Exercise:

### Data Retrieval and Understanding:

- Retrieved synthetic training dataset (synth_train.txt) and test dataset (synth_test.txt) with response variable Y in {1, 2} and feature variable X in R2.

### Data Loading and Exploration:

- Loaded training dataset into R, examined its dimension, and displayed first six records to understand size and structure.

### Data Visualization:

- Represented observations graphically, coloring points by class and modifying symbols, enhancing clarity with a legend.

### K-Nearest Neighbors (KNN) Classification:

- Utilized 'knn' function from *class* package with k = 15 neighbors to predict classes of points at (0,0) and (-2,2).

- Predicted classes of training dataset using knn with k = 10, k = 5, k = 3, and k = 1 neighbors, calculating empirical error rates.

### Decision Boundary Visualization:

- Graphically depicted decision boundary for k = 15 and k = 1 neighbors, constructing grids of points and coloring them based on predicted classes.

### Evaluation on Test Data:

- Loaded test dataset into R, predicted classes with k = 15 and k = 1 neighbors, comparing empirical error rates with training dataset predictions.

### True Positive Rate and True Negative Rate Calculation:

- Calculated true positive rate (TVP) and true negative rate (TVN) of KNN classifier with k = 15 neighbors on test dataset.

### Cost-Sensitive Classification:
Introduced cost matrix with higher cost for false negatives.
Programmed prob_knn function to estimate posterior probabilities and derived predictions using Bayes' rule with provided cost matrix.
Calculated TVP and TVN considering cost-sensitive classification.


The exercise encompassed structured tasks including data loading, exploration, visualization, KNN classification, evaluation on both training and test datasets, calculation of true positive and true negative rates, and cost-sensitive classification.
A variety of R functions and packages were utilized effectively, enhancing understanding of machine learning techniques such as KNN classification, model evaluation, and handling cost-sensitive scenarios.




