---
title: "PART II: exercise 1"
output: html_notebook
---

```{r}
data <- read.table("MCV.txt", header=TRUE)
aggregate(data[,c("AGE","CHD")], by=list(data$AGRP), FUN=mean)

plot(data$AGE, data$CHD, xlab="Age", ylab="Maladie Cardiovasculaire")
```
#Le nuage de points suggère une relation positive entre l'âge et le risque de maladie cardiovasculaire. Cependant, la relation n'est pas parfaitement linéaire, elle est logistique.
```{r}
centre <- tapply(data$AGE, data$AGRP, mean)
p <- tapply(data$CHD, data$AGRP, mean)
plot(centre, p, xlab="Centre de classe d'âge", ylab="Proportion de malades")

mod.logit <- glm(CHD ~ AGE, family=binomial(link="logit"), data=data)
summary(mod.logit)

courbe.logit <- predict(mod.logit, type="response")
plot(centre, p, ylim=c(0,1), xlab="Centre de classe d'âge", ylab="Proportion observée")
lines(sort(data$AGE), courbe.logit[order(data$AGE)], col="red")

mod.probit <- glm(CHD ~ AGE, family=binomial(link="probit"), data=data)
summary(mod.probit)
anova(mod.logit, mod.probit, test="Chisq")

newdata <- data.frame(AGE=30)
cote.logit <- predict(mod.logit, newdata, type="response")/(1-predict(mod.logit, newdata, type="response")) 
cote.probit <- predict(mod.probit, newdata, type="response")/(1-predict(mod.probit, newdata, type="response"))

exp(coef(mod.logit))
exp(coef(mod.probit))

```
# Output of the logistic regression model

The coefficient for AGE is positive and highly significant (p-value < 0.001), indicating an increased risk of cardiovascular disease with age.
The model has 98 residual degrees of freedom and an AIC of 111.5.

# Plot of observed proportions by age group
There is an increasing relationship between the proportion of diseased individuals and the center of the age group, confirming the age effect.

# Scatterplot of CHD vs AGE
The relationship is not perfectly linear, justifying the use of a logistic/probit model.

# Scatterplot of observed proportions by age group
Same information as Image 2 but in scatterplot form.

# Outputs of the logistic and probit models
The coefficients differ slightly, but the p-values are highly significant for AGE in both models.

The logistic and probit models capture the increasing relationship between age and cardiovascular disease risk similarly.
These analyses confirm a statistically significant and positive association between age and the risk of cardiovascular disease in these data. Both the logistic and probit models appear appropriate for modeling this non-linear relationship.